BAI resiliences ten droughts

### 2018-02-13 analysis/bai/trends_resilience_bai.Rmd
resiliences bais longterm

```{r packages, warning=FALSE, message=FALSE, echo=FALSE}
library("here")
library("tidyverse")
library("dplR")
library("knitr")
library('gtable')
library('grid')
library('gridExtra')
library('stringr')
library('pander')
library('broom')
library('effects')
library('devtools')
```
# Resilience

@@

machine <- 'ajpeluLap'

di <- paste0('/Users/', machine, '/Dropbox/phd/phd_repos/qpyr_dendro/', sep = '')

# sj 
sj <- read.rwl(fname=paste0(di, '/data_raw/dendro_ring/sn_sanjuan/sn_sanjuan.rwl'), format="tucson")

# canar 
ca <- read.rwl(fname=paste0(di, '/data_raw/dendro_ring/sn_canar/sn_canar.rwl'), format="tucson")

# Read diameters data
compete <- read.csv(file=paste0(di, '/data_raw/dendro_competence.csv'), header=TRUE, sep=',')
```

```{r}
source(paste0(di, 'script/R/rw_byTree.R'))
source(paste0(di, 'script/R/bai_piovesan.R'))
source(paste0(di, 'script/R/baiResilience.R'))
```

@@

```{r}
# Drought years

dy <- read.csv(file=here::here('data/sequias', 'severe_spei12.csv'), header=TRUE)

dy <- dy %>% arrange(desc(maxyear))

dys <- dy %>% filter(maxyear > 1940) %>% as.data.frame()

dys <- dys[,'maxyear']

# ojo quito 1985 
dys <- dys[which(dys!=1985)] 
```

```{r}
# SJ 
res_3_sj <- baiResilience(bai_sj, event_years = dys, window = 3)
# caL
res_3_caL <- baiResilience(bai_caL, event_years = dys, window = 3)
# caH
res_3_caH <- baiResilience(bai_caH, event_years = dys, window = 3)
```

@@

re <- bind_rows(rsj, rcaL, rcaH)
# re$disturb_year <- as.factor(re$disturb_year)
re$site <- as.factor(re$site)

re %>% ggplot(aes(x=disturb_year, ))

s <- re %>% 
  group_by(site, disturb_year) %>% 
  summarise_at(c("rt", "rs", "rc", "rrs"), funs(mean,sd,se=sd(.)/sqrt(n())))

s %>% 
  ggplot(aes(y=rs_mean, x=disturb_year, colour=site)) + 
  geom_errorbar(aes(ymin=rs_mean - rs_sd, 
                    ymax=rs_mean + rs_sd),
                width = 0.2, 
                position=position_dodge(2)) +
  geom_point(size=2, shape=21, fill='white', position=position_dodge(2)) + 
  #facet_wrap(~site, scales='fixed', ncol=1) + 
  theme_bw() # + geom_smooth(aes(colour=site), se=FALSE)

@@

>>>>

```{r packages, warning=FALSE, message=FALSE, echo=FALSE}
library("here")
library("tidyverse")
library("dplR")
library("knitr")
library('gtable')
library('grid')
library('gridExtra')
library('stringr')
library('pander')
library('broom')
library('effects')
library('devtools')
library("ggrepel")
```
# Resilience

@@

# machine <- 'ajpeluLap'

# di <- paste0('/Users/', machine, '/Dropbox/phd/phd_repos/qpyr_dendro/', sep = '')

# sj 
sj <- read.rwl(fname=here::here('data_raw/dendro_ring/sn_sanjuan', 'sn_sanjuan.rwl'), format="tucson")

# canar 
ca <- read.rwl(fname=here::here('data_raw/dendro_ring/sn_canar', 'sn_canar.rwl'), format="tucson")

# Read diameters data
compete <- read.csv(file=here::here('data_raw', 'dendro_competence.csv'), header=TRUE, sep=',')
```

```{r}
source(here::here('script/R','rw_byTree.R'))
source(here::here('script/R', 'bai_piovesan.R'))
source(here::here('script/R', 'baiResilience.R'))
```

@@

```{r}
# Drought years 
# Utilizamos spei12 
dy <- read.csv(file=here::here('data/sequias', 'severe_spei12.csv'), header=TRUE)
dy <- dy %>% arrange(desc(maxyear))

# solamente aquellos mayor de 2 meses 
dys <- dy %>% filter(d_duration > 2) %>% as.data.frame() 
dys <- dys[,'maxyear']

```{r}
# Computar resiliencia para otros years 
# SJ 
res_3_sj <- baiResilience(bai_sj, event_years = dys, window = 3)
# Remove 1914 
dyssj <- dys[which(dys!=1914)]
res10sj <- baiResilience(bai_sj, event_years = dyssj, window = 10)

# caL
res_3_caL <- baiResilience(bai_caL, event_years = dys, window = 3)
res10caL <- baiResilience(bai_caL, event_years = dys, window = 10)
# caH
res_3_caH <- baiResilience(bai_caH, event_years = dys, window = 3)
res10caH <- baiResilience(bai_caH, event_years = dys, window = 10)
```

@@

rsj10 <- res10sj$resilience %>% mutate(site='SJ')
rcaL10<- res10caL$resilience %>% mutate(site='caL')
rcaH10 <- res10caH$resilience %>% mutate(site='caH')

re <- bind_rows(rsj, rcaL, rcaH)
re10 <- bind_rows(rsj10, rcaL10, rcaH10)
# re$disturb_year <- as.factor(re$disturb_year)
re$site <- as.factor(re$site)
re10$site <- as.factor(re10$site)

## Sample depth (#trees by site)
sampleDepth <- function(x, site){ 
  x$samps <- rowSums(!is.na(x))
  x$year <- as.numeric(row.names(x))
  x$site <- site
  out <- x[,c("samps", "year", "site")]
  return(out)
}

d_sj <- sampleDepth(sj_tree, site='SJ')
d_caL <- sampleDepth(caL_tree, site='caL')
d_caH <- sampleDepth(caH_tree, site='caH')

d <- rbind(d_sj, d_caL, d_caH)

# Get mean values by site and add sample deepth
s <- re %>% 
  group_by(site, disturb_year) %>% 
  summarise_at(c("rt", "rs", "rc", "rrs"), funs(mean(., na.rm = TRUE), sd(., na.rm=TRUE),se=sd(., na.rm=TRUE)/sqrt(n()))) %>% 
  inner_join(d, by=c("disturb_year" = "year", "site"="site")) %>% 
  inner_join(dy, by=c("disturb_year" = "maxyear"))

s10 <- re10 %>% 
  group_by(site, disturb_year) %>% 
  summarise_at(c("rt", "rs", "rc", "rrs"), funs(mean(., na.rm = TRUE), sd(., na.rm=TRUE),se=sd(., na.rm=TRUE)/sqrt(n()))) %>% 
  inner_join(d, by=c("disturb_year" = "year", "site"="site")) %>% 
  inner_join(dy, by=c("disturb_year" = "maxyear"))

@@

```{r}
# ojo he ploteado los que tiene mas de 10 árboles 
resi_temporal <- s %>% filter(samps >= 10) %>% 
  ggplot(aes(y=rs_mean, x=disturb_year, colour=site)) + 
  geom_errorbar(aes(ymin=rs_mean - rs_sd, 
                    ymax=rs_mean + rs_sd),
                width = 0.2, 
                position=position_dodge(2)) +
  geom_point(aes(size=d_duration), shape=21, fill='white', position=position_dodge(2)) + 
  facet_wrap(~site, scales='free_y', ncol=1) + 
  theme_bw() + 
  theme(strip.background = element_rect(fill='transparent')) + 
  ylab('Resilience (Rs)') + xlab('years') + 
  scale_x_continuous(breaks=seq(1910,2020, by=5)) # + geom_smooth(aes(colour=site), se=FALSE)

resi_temporal

pdf(here::here('/out/bai_resiliences_severes', 'res_bai_longterm.pdf'), width=12, height = 9)
resi_temporal
dev.off()

resi_temporal10 <- s10 %>% filter(samps >= 10) %>% 
  ggplot(aes(y=rs_mean, x=disturb_year, colour=site)) + 
  geom_errorbar(aes(ymin=rs_mean - rs_sd, 
                    ymax=rs_mean + rs_sd),
                width = 0.2, 
                position=position_dodge(2)) +
  geom_point(aes(size=d_duration), shape=21, fill='white', position=position_dodge(2)) + 
  facet_wrap(~site, scales='free_y', ncol=1) + 
  theme_bw() + 
  theme(strip.background = element_rect(fill='transparent')) + 
  ylab('Resilience (Rs)') + xlab('years') + 
  scale_x_continuous(breaks=seq(1910,2020, by=5)) # + geom_smooth(aes(colour=site), se=FALSE)

pdf(here::here('/out/bai_resiliences_severes', 'res_bai_longterm10.pdf'), width=12, height = 9)
resi_temporal10
dev.off()

```{r}

#https://gist.github.com/ottadini/6882677 
lm_eqn = function(m) {
  # Displays regression line equation and R^2 value on plot
  # Usage:
  # p + annotate("text", x=25, y=300, label=lm_eqn(lm(y ~ x, df)), parse=TRUE)

  l <- list(#a = format(coef(m)[1], digits = 2),
            # b = format(abs(coef(m)[2]), digits = 2),
            r2 = format(summary(m)$r.squared, digits = 3));

  if (coef(m)[2] >= 0)  {
    # eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,l)
    eq <- substitute(~~italic(r)^2~"="~r2,l)
  } else {
    # eq <- substitute(italic(y) == a - b %.% italic(x)*","~~italic(r)^2~"="~r2,l) 
    eq <- substitute(~~italic(r)^2~"="~r2,l)
  }

  as.character(as.expression(eq));                 
}

sdf <- s %>% filter(samps >= 10) %>% group_by(disturb_year) %>% 
  summarise(rs_avg = mean(rs_mean, na.rm=TRUE), 
            rt_avg = mean(rt_mean, na.rm=TRUE), 
            rc_avg = mean(rc_mean, na.rm=TRUE)) %>% 
  inner_join(dy, by=c("disturb_year" = "maxyear"))

tR <- list(geom_smooth(method='lm', se=FALSE),
           theme_bw(), xlab("Drought Severity"), 
          # scale_size_continuous("Drought Intensity", breaks = c(-2,-1.8,-1.6, -1.4, -1.2), labels= c(-2,-1.8,-1.6, -1.4, -1.2)),
          geom_text_repel(),
          scale_size_continuous("Drought Intensity (*-1)"),
          geom_point(),
          theme(legend.position ='top'))

sevRs <- sdf %>% 
  ggplot(aes(y=rs_avg,  x=d_severity, label=disturb_year)) +  
  geom_point(aes(size=-1*d_intensity)) + 
  ylab("Resilience (Rs) (3year window size)") + tR +
  annotate("text", x=7, y = 1.4, label=lm_eqn(lm(rs_avg~d_severity, sdf)), parse=TRUE) 
sevRs

sevRt <- sdf %>% 
  ggplot(aes(y=rt_avg,  x=d_severity, label=disturb_year)) + 
  geom_point(aes(size=-1*d_intensity)) + 
  ylab("Resistance (Rt) (3year window size)") + tR +
  annotate("text", x=7, y = 1.3, label=lm_eqn(lm(rt_avg~d_severity, sdf)), parse=TRUE) 
sevRt

sevRc <- sdf %>% 
  ggplot(aes(y=rc_avg,  x=d_severity, label=disturb_year)) + 
  geom_point(aes(size=-1*d_intensity)) + 
  ylab("Recovery (Rc) (3year window size)") + tR +
  annotate("text", x=7, y = 2, label=lm_eqn(lm(rc_avg~d_severity, sdf)), parse=TRUE) 
sevRc

pdf(here::here('/out/bai_resiliences_severes', 'rs_vs_drought_3years.pdf'), width=18, height = 7)
sevRt + sevRc + sevRs 
dev.off()

sdf10 <- s10 %>% filter(samps >= 10) %>% group_by(disturb_year) %>% 
  summarise(rs_avg = mean(rs_mean, na.rm=TRUE), 
            rt_avg = mean(rt_mean, na.rm=TRUE), 
            rc_avg = mean(rc_mean, na.rm=TRUE)) %>% 
  inner_join(dy, by=c("disturb_year" = "maxyear"))

sevRs10 <- sdf10 %>% 
  ggplot(aes(y=rs_avg,  x=d_severity, label=disturb_year)) +  
  geom_point(aes(size=-1*d_intensity)) + 
  ylab("Resilience (Rs) (10year window size)") + tR +
  annotate("text", x=7, y = 1.4, label=lm_eqn(lm(rs_avg~d_severity, sdf10)), parse=TRUE) 
sevRs10

sevRt10 <- sdf10 %>% 
  ggplot(aes(y=rt_avg,  x=d_severity, label=disturb_year)) + 
  geom_point(aes(size=-1*d_intensity)) + 
  ylab("Resistance (Rt) (10year window size)") + tR +
  annotate("text", x=7, y = 1.3, label=lm_eqn(lm(rt_avg~d_severity, sdf10)), parse=TRUE) 
sevRt10

sevRc10 <- sdf10 %>% 
  ggplot(aes(y=rc_avg,  x=d_severity, label=disturb_year)) + 
  geom_point(aes(size=-1*d_intensity)) + 
  ylab("Recovery (Rc) (10year window size)") + tR +
  annotate("text", x=7, y = 2, label=lm_eqn(lm(rc_avg~d_severity, sdf10)), parse=TRUE) 
sevRc10

pdf(here::here('/out/bai_resiliences_severes', 'rs_vs_drought_10years.pdf'), width=18, height = 7)
sevRt10 + sevRc10 + sevRs10 
dev.off()
```
